---
title: "Movies_Recommendation_Engine"
author: "S.Vijay"
date: "29 September 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("E:/Kaggle/Movies Dataset")

library(data.table)
library(dplyr)
library(stringr)
library(jsonlite)
library(tidyverse)
library(reshape)
library(tm)
library(RSentiment)
library(wordcloud)
library(ggplot2)
library(corrplot)
library(caret)
library(randomForest)
library(reshape2)
library(recommenderlab)

```

```{r}
movies = read.csv("movies_metadata.csv", stringsAsFactors = F)
rating = read.csv("ratings_small.csv")
```

#### Summary of Dataset
```{r}
glimpse(movies)
glimpse(rating)
```

### EDA
#### Check for NA values
```{r}
colSums(is.na(movies))
```

#### Removing the NA values
```{r}
movies_df = na.omit(movies)
```

#### Final check for NA values
```{r}
colSums(is.na(movies_df))
```

#### Creating new dataframe for genre column in movies dataset which is in JSON format
```{r}
movies_df$genres = gsub("'","\"", movies_df$genres)

genre_df = movies_df %>% filter(nchar(genres) > 2) %>% 
                      mutate(js = lapply(genres, fromJSON)) %>% 
                      unnest(js) %>% select(id, title, genre = name)

genre_bud_rev = movies_df %>% select(id, budget, revenue)

genre = merge(genre_df, genre_bud_rev, by.x = "id", by.y = "id", all.x = T)

movie_genre = genre %>% filter(budget != 0 & revenue != 0)
```

#### Collabarative Filtering using Movie Genre
```{r}
m = cast(movie_genre, formula = title~genre)
colfil = merge(m, movies_df, by.x = "title", by.y = "title", all.x = T)
colfil_df = colfil[,c(1:21,23,29,31,36,37)]
```

#### EDA
```{r}
# Deriving Difference column
movies_df$Difference = ifelse(movies_df$budget == 0 | movies_df$revenue == 0, 
                              0, movies_df$revenue - movies_df$budget)

# Deriving Profit percent column
movies_df$profit_percent = ((movies_df$Difference*100)/movies_df$budget)

movies_df$Result = ifelse(movies_df$profit_percent >= 80,"Block_Buster",
                          ifelse(movies_df$profit_percent >= 50 & movies_df$profit_percent < 80, "Super_Hit",
                                 ifelse(movies_df$profit_percent >= 30 & movies_df$profit_percent < 50, "Hit",
                                        ifelse(movies_df$profit_percent > 0 & movies_df$profit_percent < 30, "Average_Hit",
                                               ifelse(movies_df$profit_percent == 0, "Neutral","Flop")))))

movies_df$Result = ifelse(is.na(movies_df$Result),"Neutral",movies_df$Result)

# Converting character column to catagorical column
movies_df$original_language = as.factor(movies_df$original_language)
movies_df$original_title = as.factor(movies_df$original_title)
movies_df$status = as.factor(movies_df$status)
movies_df$Result = as.factor(movies_df$Result)

# Sparating numeric and categorical column
numeric_col <- movies_df[sapply(movies_df, is.numeric)]

categorical_col <- movies_df[sapply(movies_df,is.factor)]

```

#### Correlation Analysis
```{r}
corr <- cor(movies_df[sapply(movies_df, is.numeric)])
corrplot(corr,type = "upper",method = "number") # Plot correlation matrix
corr[lower.tri(corr)] = 0
corr <- corr-diag(nrow(corr))
index <- which(abs(corr) > 0.7,arr.ind = T)
df <- data.frame(rowname=rownames(index),colname=colnames(corr)[index[,2]],cor_value=corr[index])
df
```

### Univariate Analysis on Categorical columns - Checking for distribution
```{r}
for(i in 1:ncol(categorical_col)){
  g = plot(categorical_col[,i], main = paste("Bar plot of", names(categorical_col[i])), xlab = names(categorical_col[i]))
}
```


### Univariate Analysis on Numeric columns - Checking for distribution & skewness
```{r}
for(i in 1:ncol(numeric_col)){
  g = hist.default(numeric_col[,i], main = paste("Histogram of", names(numeric_col[i])), xlab = names(numeric_col[i]))
}
```

#### Exploring the tagline column to gather information on most used tagwords
```{r}

# taking the tagline column and separating each words
tagline = gsub("[^A-Za-z///]", " ", movies_df$tagline)

# Creating the corpus for text mining
corpus = Corpus(VectorSource(list(movies_df$tagline))) 

# Removing punctuation from the corpus
corpus = tm_map(corpus,removePunctuation)

# Converting uppercase letters to lowecase
corpus = tm_map(corpus,content_transformer(tolower))

# Removing the unwanted whitespaces
corpus = tm_map(corpus,stripWhitespace)

# Removing the stopwords from the corpus
corpus = tm_map(corpus,removeWords,stopwords("english"))

# Finally creating the document term matrix for doing sentiment analysis
dtm_tag = DocumentTermMatrix(VCorpus(VectorSource(corpus[1]$content)))

freq_tag=colSums(as.matrix(dtm_tag))

sent_tag=calculate_sentiment(names(freq_tag)) %>% cbind(as.data.frame(freq_tag))

positive=sent_tag[sent_tag$sentiment=="Positive",]
negative=sent_tag[sent_tag$sentiment=="Negative",]
cat("Positive Sentiment:",sum(positive$freq_tag),"Negative Sentiment:",sum(negative$freq_tag),sep="\n")
```

#### Word cloud for positive and negative tagwords 
```{r,warning=FALSE}
#par(mfrow=c(1,2))
wordcloud(positive$text,positive$freq_tag, min.freq=10,rot.per=0,vfont=c("sans serif","plain"))
text(x=0.5, y=1.05, "Positive",col=brewer.pal(4, "Reds"))
wordcloud(negative$text,negative$freq_tag, min.freq=7,rot.per=0,vfont=c("sans serif","plain"))
text(x=0.5, y=1.05, "Negative",col=brewer.pal(4, "Reds"))
```

#### Which year has seen maximum release of movies?
```{r}
movies_df$release_date = as.Date(movies_df$release_date, format = "%d-%m-%Y")

movies_df$Year=as.factor(format(movies_df$release_date,"%Y"))
movies_df$Date=as.factor(format(movies_df$release_date,"%d"))
movies_df$month=month.abb[(as.factor(format(movies_df$release_date,"%m")))]

movies_df %>% group_by(month) %>% drop_na(month) %>% summarise(count=n()) %>% arrange(desc(month)) %>% ggplot(aes(reorder(month,count),count,fill=month))+geom_bar(stat="identity")+theme_classic()+theme(plot.title=element_text(size=14,face="italic",colour="red"),legend.position="none")+labs(x="",y="",title="Number of Movies Released per month")+coord_flip()+geom_label(aes(label=count))


```

#### Highest Budget& Revenue Movies

```{r}
movies_df %>% select(original_title,budget) %>% 
              drop_na(original_title) %>% 
              arrange(desc(budget)) %>% 
              head(10) %>% 
              ggplot(aes(reorder(original_title,budget), budget,fill=original_title)) + 
              geom_bar(stat="identity") + theme_classic() +
              theme(axis.text.x = element_text(angle=90), 
                    plot.title=element_text(color="Red",face="italic"),
                    legend.position="none") + scale_y_continuous(labels=scales::comma) + 
              labs(x="",y="Total Budget in $",title="Most Expensive Movies -Top 10")
```


#### Highest Grossing Movies

```{r}
movies_df %>% select(original_title,revenue) %>% drop_na(original_title) %>% 
              arrange(desc(revenue)) %>% 
              head(10) %>% 
              ggplot(aes(reorder(original_title,revenue),revenue,fill=original_title)) + 
              geom_bar(stat="identity") + theme_classic() + 
              theme(axis.text.x=element_text(angle=90),plot.title=element_text(color="Red",face="italic"),
                    legend.position="none") + scale_y_continuous(limits=c(0,3000000000),
                                                                 breaks=seq(0,3000000000,500000000),
                                                                 labels=scales::comma) + 
              labs(x="",y="Total Revenue in $",title="Highest Grossing Movies -Top 10")
```


#### Movies with Highest Popularity
```{r}
movies_df %>% group_by(original_title) %>% 
              arrange(desc(popularity)) %>% 
              head(10) %>% 
              ggplot(aes(factor(original_title,levels=original_title),
                         popularity,fill=original_title)) + 
              geom_bar(stat="identity") + 
              theme_classic() + 
              theme(axis.text.x=element_text(angle=90,hjust=0.5),
                    plot.title=element_text(hjust=0.5,size=15,color="red"),
                    legend.position="none") + 
              labs(x="Title",y="Rating",title="Popularity of Movies") + 
              scale_x_discrete(labels=function(x) str_wrap(x,width=15))
```

#### Movies with Highest Votes
```{r}
movies_df %>% group_by(original_title) %>% arrange(desc(vote_count)) %>% head(10) %>% ggplot(aes(factor(original_title,levels=original_title),vote_count,fill=original_title))+geom_bar(stat="identity")+theme_classic()+theme(axis.text.x=element_text(angle=90,hjust=0.5),plot.title=element_text(hjust=0.5,size=15,color="red"),legend.position="none")+labs(x="Title",y="Vote Count",title="Vote Count of Movies")+scale_x_discrete(labels=function(x) str_wrap(x,width=15))
```

#### Predictive Model for Revenue of the movies

```{r}
# Preparing train and test data for model building
mov_train = movies_df[sample(1:nrow(movies_df),0.8*nrow(movies_df)),]
mov_test = movies_df[sample(1:nrow(movies_df),0.2*nrow(movies_df)),]


revenue_model_1 = lm(revenue~budget, data = mov_train) # Model 1 with only budget

revenue_model_2 = lm(revenue~budget+vote_count, data = mov_train) # Model 2 with budget and vote_count

revenue_model_3 = lm(revenue~budget+Difference, data = mov_train) # Model 3 with budget and difference

revenue_model_4 = lm(revenue~budget+vote_count+Difference, data = mov_train) # Model 4 with budget, vote_count and difference


# Summaray of all the models to find which model performs well
summary(revenue_model_1)
summary(revenue_model_2)
summary(revenue_model_3)
summary(revenue_model_4)
```

#### Revenue Predictions
```{r}
mov_test$predicted_Result_1 = predict(revenue_model_1, mov_test %>% select(-revenue))
mov_test$predicted_Result_2 = predict(revenue_model_2, mov_test %>% select(-revenue))
mov_test$predicted_Result_3 = predict(revenue_model_3, mov_test %>% select(-revenue))
mov_test$predicted_Result_4 = predict(revenue_model_4, mov_test %>% select(-revenue))
```


#### Checking the RMSE value for better performing model
```{r}
rmse_1 = sqrt(mean(mov_test$revenue - mov_test$predicted_Result_1)^2)

rmse_2 = sqrt(mean(mov_test$revenue - mov_test$predicted_Result_2)^2)

rmse_3 = sqrt(mean(mov_test$revenue - mov_test$predicted_Result_3)^2)

rmse_4 = sqrt(mean(mov_test$revenue - mov_test$predicted_Result_4)^2)

rmse = c(rmse_1,rmse_2,rmse_3,rmse_4)
lm_model = c("revenue_model_1","revenue_model_2","revenue_model_3","revenue_model_4")

rmse_df = data.frame(rmse, lm_model)

rmse_df %>% arrange(rmse_df$rmse) %>% head(1) # Best performing model = revenue_model_3

```

#### Classification model for finding the result of the movies
```{r}
movies_df$profit_percent = ifelse(is.na(movies_df$profit_percent), 0,movies_df$profit_percent)

data = movies_df %>% mutate_if(is.character, as.factor) # Converting character column to factor column 

data$release_date = as.Date(data$release_date,format = "%d-%m-%Y") 

data$release_date = ifelse(is.na(data$release_date), 1,data$release_date) # missing values in date column is treated as 1
```


#### Random Forest Classification Model
```{r}
# Test and train data
rf_train = data[sample(1:nrow(data),0.7*nrow(data)),]
rf_test = data[sample(1:nrow(data),0.3*nrow(data)),]

mtry = round(sqrt(length(colnames(rf_train))-1)) # Sampling the columns of train data

model_rf = randomForest(Result~., data = rf_train %>% select(-original_language,
                                                             -original_title,
                                                             -tagline,
                                                             -title,
                                                             -overview,
                                                             -genres), 
                        ntree = 100, mtry = mtry) # Model building

rf_test$Predicted_Result_1 = predict(model_rf, rf_test) # Predicting the result

cm = confusionMatrix(rf_test$Predicted_Result_1, rf_test$Result) # Checking the model performance
cm
cm$overall['Accuracy']*100

```

#### Collabrative filtering for Movie Genre
```{r}
MoviesUmayLIKE = function(x){
if(length(Genre) == 1 & any(Genre %in% colnames(colfil_df))){
  Recommended_Movies = colfil_df %>% filter(colfil_df[Genre] == 1) %>% arrange(-popularity) %>% head(10) %>%  
                                select("title","popularity","budget","revenue")
  }else if(length(Genre) == 2 & any(Genre %in% colnames(colfil_df))){
    Recommended_Movies = colfil_df %>% filter(colfil_df[Genre][1] == 1 & colfil_df[Genre][2] == 1) %>% 
      arrange(-popularity) %>% head(10) %>% select("title","popularity","budget","revenue")
    }else if(length(Genre) == 3 & any(Genre %in% colnames(colfil_df))){
      Recommended_Movies = colfil_df %>% filter(colfil_df[Genre][1] == 1 & colfil_df[Genre][2] == 1 & 
                                                  colfil_df[Genre][3] == 1) %>% arrange(-popularity) %>% 
        head(10) %>% select("title","popularity","budget","revenue")
      }else if(length(Genre) == 4 & any(Genre %in% colnames(colfil_df))){
        Recommended_Movies = colfil_df %>% filter(colfil_df[Genre][1] == 1 & colfil_df[Genre][2] == 1 & 
                                                    colfil_df[Genre][3] == 1 & colfil_df[Genre][4] == 1) %>% 
          arrange(-popularity) %>% head(10) %>% select("title","popularity","budget","revenue")
        }else if(length(Genre) == 5 & any(Genre %in% colnames(colfil_df))){
          Recommended_Movies = colfil_df %>% filter(colfil_df[Genre][1] == 1 & colfil_df[Genre][2] == 1 & 
                                                      colfil_df[Genre][3] == 1 & colfil_df[Genre][4] == 1 & 
                                                      colfil_df[Genre][5] == 1) %>% arrange(-popularity) %>% 
            head(10) %>% select("title","popularity","budget","revenue")
          }else{
    print("Genre not found in our database")
          }
 return(Recommended_Movies)
}
```

#### Getting input from the user to filter movies accordingly
```{r}
#n = as.integer(readline(prompt = "Enter the number of genre you want to filter not greater than 5: "))
#Genre = c()
#for (i in 1:n) {
#  Genre[i] = readline(prompt = "Enter Movie genre: ")
#  print(Genre)
#}
#MoviesUmayLIKE(Genre) # Function to get the output
```

#### Recommendation System
```{r}
length(unique(rating$userId)) # Totally 671 users
length(unique(rating$movieId)) # Totally 9066 movies
```

#### User Based Collabrative Filtering
```{r}
# Creating the rating matrix with rows as uers and columns as movies
ratings_matrix = as.matrix(dcast(data = rating, userId~movieId, value.var = "rating"))

# We can now remove the user ids and convert ratings_matrix to real rating matrix
ranking_matrix = as(ratings_matrix[,-1], 'realRatingMatrix')

# Create Recommender Model.The parameters are UBCF and Cosine similarity. We take 10 nearest neighbours
ubcf = Recommender(ranking_matrix, method = 'UBCF', param = list(method = 'Cosine', nn = 10))

# Predicting the recommendation of 4th user
result = predict(ubcf, ranking_matrix[4, ])

# Top 5 movies recommended to 4th user
movies_sugg = as((bestN(result, n=5)),"list")[[1]]

# Output with movieid and title
movies %>% filter(id %in% movies_sugg) %>% select(id,title)

```

#### Item Based Collabrative Filtering
```{r}
# Creating the rating matrix with rows as movies and columns as users
ratings_matrix_1 = as.matrix(dcast(data = rating, movieId~userId, value.var = 'rating'))

# We can now remove the movie ids and convert ratings_matrix to real rating matrix
ranking_matrix_1 = as(ratings_matrix[,-1], 'realRatingMatrix')

# Create Recommender Model.The parameters are IBCF and Cosine similarity. We take 10 nearest neighbours
ibcf = Recommender(ranking_matrix_1, method = "IBCF", param = list(method = "Cosine", k = 10))

# Predicting the recommendation of 4th user
result_1 = predict(ibcf, ranking_matrix_1[4, ])

# Top 5 movies recommended to 4th user
movies_sugg_1 = as((bestN(result_1, n=5)), "list")[[1]]

# Output with movieid and title
movies %>% filter(movieId %in% movies_sugg_1) %>% select(movieId,title)

```
#### Evaluation of IBCF and UBCF model to find the best performing model
```{r}
# create evaluation scheme splitting taking 90% of the date for training and leaving 10% for validation or test
u1 <- evaluationScheme(ranking_matrix, method="split", train=0.9, given=-1)
u2 <- evaluationScheme(ranking_matrix, method="split", train=0.8, given=-1)
u3 <- evaluationScheme(ranking_matrix, method="split", train=0.7, given=-1)

i1 <- evaluationScheme(ranking_matrix_1, method="split", train=0.9, given=-1)
i2 <- evaluationScheme(ranking_matrix_1, method="split", train=0.8, given=-1)
i3 <- evaluationScheme(ranking_matrix_1, method="split", train=0.7, given=-1)

# creation of recommender model based on ubcf
Rec.ubcf_1 <- Recommender(getData(u1, "train"), "UBCF")
Rec.ubcf_2 <- Recommender(getData(u2, "train"), "UBCF")
Rec.ubcf_3 <- Recommender(getData(u3, "train"), "UBCF")

# creation of recommender model based on ibcf for comparison
Rec.ibcf_1 <- Recommender(getData(i1, "train"), "IBCF")
Rec.ibcf_2 <- Recommender(getData(i2, "train"), "IBCF")
Rec.ibcf_3 <- Recommender(getData(i3, "train"), "IBCF")

# making predictions on the test data set
p.ubcf_1 <- predict(Rec.ubcf_1, getData(u1, "known"), type="ratings")
p.ubcf_2 <- predict(Rec.ubcf_2, getData(u2, "known"), type="ratings")
p.ubcf_3 <- predict(Rec.ubcf_3, getData(u3, "known"), type="ratings")

# making predictions on the test data set
p.ibcf_1 <- predict(Rec.ibcf_1, getData(i1, "known"), type="ratings")
p.ibcf_2 <- predict(Rec.ibcf_2, getData(i2, "known"), type="ratings")
p.ibcf_3 <- predict(Rec.ibcf_3, getData(i3, "known"), type="ratings")

# obtaining the error metrics for both approaches and comparing them
error.ubcf_1<-calcPredictionAccuracy(p.ubcf_1, getData(u1, "unknown"))
error.ubcf_2<-calcPredictionAccuracy(p.ubcf_2, getData(u2, "unknown"))
error.ubcf_3<-calcPredictionAccuracy(p.ubcf_3, getData(u3, "unknown"))

error.ibcf_1<-calcPredictionAccuracy(p.ibcf_1, getData(i1, "unknown"))
error.ibcf_2<-calcPredictionAccuracy(p.ibcf_2, getData(i2, "unknown"))
error.ibcf_3<-calcPredictionAccuracy(p.ibcf_3, getData(i3, "unknown"))
error <- rbind(error.ubcf_1,error.ubcf_2,error.ubcf_3,error.ibcf_1,error.ibcf_2,error.ibcf_3)
rownames(error) <- c("model_1_UBCF","model_2_UBCF","model_3_UBCF","model_1_IBCF","model_2_IBCF","model_3_IBCF")
error
```

